---
title: "Location"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: location
listpages: false
weight: 300
summary: "Civil servants by location"
---

While popular perception is that civil servants all work in Whitehall, only one in five civil servants (20.9%) work in London, civil servants work in over 2,500 individual locations across the UK.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

data_dir <- govukhugo::data_dir()

acses_postcodes <- read_csv(
  file.path(data_dir, "acses2021_postcodes.csv")
)

acses_nspl <- read_csv(
  file.path(data_dir, "acses_nspl.csv")
) %>%
  sf::st_as_sf(
    coords = c("long", "lat"),
    crs = sf::st_crs(4326))

acses_postcode_geo <- left_join(
  acses_postcodes,
  acses_nspl,
  by = "postcode"
)

uk_countries_json <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2021_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"

uk_background <- sf::st_read(uk_countries_json)

all_emp_postcode <- acses_postcode_geo %>%
  filter(organisation == "All employees" & 
           responsibility_level == "All employees" & !is.na(pcds) &
           value != 0) %>%
  mutate(fill_value = log10(value)) %>%
  arrange(value)

```


::::{.govuk-grid-row}
:::{.govuk-grid-column-one-half}
In this section you can explore statistics about where civil servants work:

- civil servants by [country and region](country_region/),
- civil servants by [local areas](nuts3/)
- regional distribution by [organisation and grade](organisation_grade/)
- regional distribution by [profession and function](profession_function/)
- regional distribution by [key diversity characteristics](diversity/)
:::

:::{.govuk-grid-column-one-half}
```{r postcode_chart}

postcode_plot <- ggplot(uk_background) +
  geom_sf(
    fill = govukhugo::govuk_colours["mid-grey"], 
    colour = govukhugo::govuk_colours["mid-grey"],
    size = 0.2
  ) +
  geom_sf(
    data = all_emp_postcode, 
    aes(fill = fill_value, geometry = geometry),
    shape = 21,
    colour = govukhugo::govuk_colours["chart-white"],
    stroke = 0.15,
    alpha = 0.5,
    size = 1.5,
    show.legend = FALSE
  ) +
  scale_fill_gradientn(colours = govukhugo::govuk_palette("blues")[3:1]) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = govukhugo::govuk_colours["light-grey"],
      colour = NA
    )
  )

govukhugo::render_svg(
  postcode_plot,
  width = 300,
  height = 400,
  alt_title = "Map of Civil Service locations in the UK",
  alt_desc = paste(
    "An outline map of the United Kingdom with points showing all the",
    "postcode locations where civil servants are recorded as working as",
    "at 31 March 2021."
  )
)

```
:::
::::
