---
title: "Pay by grade and organisation"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: pay
weight: 200
summary: "Headline statistics on pay by grade and organisation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

gbp_format <- scales::dollar_format(accuracy = 1, prefix = "£")

grade_org_salary <- acses2021 %>%
  filter(source_table == "table_25") %>%
  select(responsibility_level, organisation,value_type, value) %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7", 
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants", "All employees")
  )

median_salary <- grade_org_salary %>%
  filter(organisation == "All employees" & responsibility_level == "All employees")

grade_salary <- grade_org_salary %>%
  filter(organisation == "All employees" & responsibility_level != "All employees") %>%
  mutate(
    chart_label = paste0(responsibility_level, " (", gbp_format(value), ")"),
    chart_label = fct_rev(fct_reorder(chart_label, as.numeric(responsibility_level)))
  )

org_salary <- grade_org_salary %>%
  filter(organisation != "All employees" & responsibility_level == "All employees") %>%
  left_join(group_lookup, by = "organisation")

```

As at 31 March 2021, the median salary by responsibility level ranged from `r gbp_format(min(grade_salary$value))` for `r grade_salary$responsibility_level[grade_salary$value == min(grade_salary$value)]` to `r gbp_format(max(grade_salary$value))` for `r grade_salary$responsibility_level[grade_salary$value == max(grade_salary$value)]`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="grade-salary-chart">}}
```{r grade_salary_chart}
grade_salary_plot <- ggplot(grade_salary, aes(x = chart_label, y = value)) +
  geom_col(fill = govukhugo::govuk_colours["blue"]) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1, margin = margin(r = 3)),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  grade_salary_plot,
  width = 800,
  height = 125,
  alt_title = "Median salary by responsibility level",
  alt_desc = paste(
    "A bar chart showing the median salary for each responsibility level",
    "ordered by seniority with the most senior responsibility level at the top",
    "and the most junior responsibility level at the bottom."
  ),
  caption = "alt_title"
)
```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="grade-salary-table">}}
```{r grade_salary_table}
grade_salary_table <- grade_salary %>%
  select(responsibility_level, value) %>%
  arrange(responsibility_level)

govukhugo::govuk_datatable(
    grade_salary_table,
    title = "Median salary by responsibility level",
    col_names = c("Responsibilty level", "Median salary"),
    page_length = 10,
    search = FALSE
  ) %>%
  DT::formatCurrency(columns = 2, currency = "£", digits = 0)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

As at 31 March 2021, the median salary by organisation ranged from `r gbp_format(min(org_salary$value))` for civil servants based at `r org_salary$organisation[org_salary$value == min(org_salary$value)]` to `r gbp_format(max(org_salary$value))` for civil servants based at `r org_salary$organisation[org_salary$value == max(org_salary$value)]`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="dept-salary-chart">}}
```{r dept_salary_chart}
dept_salary_plot <- ggplot(org_salary, aes(x = fct_rev(dept_group), y = value)) +
  geom_point(colour = govukhugo::govuk_colours["blue"]) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1, margin = margin(r = 3)),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    panel.grid.major.y = element_line(colour = govukhugo::govuk_colours["mid-grey"])
  )

govukhugo::render_svg(
  dept_salary_plot,
  width = 800,
  height = 600,
  alt_title = "Median salary by departmental group/organisation",
  alt_desc = paste(
    "A bar chart showing the median salary for each organisation grouped by",
    "their departmental group (in alphabetical order)."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::
