---
title: "Civil servants by salary band"
date: 2021-01-01
section: pay
weight: 100
summary: "Headline statistics on civil servants by salary band"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

gbp_format <- scales::dollar_format(accuracy = 1, prefix = "£")

salary_band_all <- acses2021 %>%
  filter(source_table == "table_06" & 
           working_pattern == "All employees" &
           sex == "All employees") %>%
  select(salary_band, value) %>%
  mutate(valid = if_else(salary_band == "Not reported" | salary_band == "All employees", NA_real_, value),
         pc = valid/sum(valid, na.rm = TRUE),
         cpc = cumsum(pc),
         icpc = 1-cpc,
         band_lower = as.numeric(
           str_remove_all(
             str_replace(salary_band, "^[£](\\d+),(\\d+).*$", "\\1\\2"),
             "\\D")
         ),
         band_upper = as.numeric(
           str_remove_all(
             str_replace(salary_band, ".* - (\\d+),(\\d+)", "\\1\\2"),
             "\\D")
         ),
         salary_band = fct_reorder(salary_band, band_lower),
         salary_band = fct_relevel(salary_band, "Not reported", "All employees", after = 0L)
         )

salary_stats <- acses2021 %>%
  filter(source_table == "table_06a" & 
           working_pattern == "All employees" &
           sex == "All employees") %>%
  select(salary_band, value_type, value)

salary_stats_out <- purrr::set_names(
  x  = gbp_format(salary_stats$value),
  nm = salary_stats$value_type
)

lowest_50 <- salary_band_all %>%
  filter(cpc >= 0.5 & icpc <= 0.5) %>%
  slice_min(cpc)

top_10 <- salary_band_all %>%
  filter(cpc >= 0.9 & icpc <= 0.1) %>%
  slice_min(cpc)

high_earners <- acses2021 %>%
  filter(source_table == "table_35") %>%
  select(salary_band, value) %>%
  arrange(value)

```

As at 31 March 2021, and `r scales::percent(lowest_50$cpc, 0.1)` of civil servants earnt `r gbp_format(lowest_50$band_upper)` or less, while `r scales::percent(top_10$icpc, 0.1)` earnt `r gbp_format(top_10$band_lower)` or more. There are `r scales::comma(high_earners$value[1], 1)` civil servants who earnt more than £100,000, of these `r scales::comma(high_earners$value[2], 1)` earnt more than £150,000 and `r scales::comma(high_earners$value[3], 1)` who earnt more than £200,000.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="salary-chart">}}
```{r salary_chart}

salary_chart_df <- salary_band_all %>%
  filter(salary_band != "All employees") %>%
  select(salary_band, value) %>%
  mutate(
    chart_label = paste0(salary_band, " (", scales::comma(value), ")"),
    fill_group = if_else(salary_band == "Not reported", "Not reported", "Reported"),
    chart_label = fct_reorder(chart_label, as.numeric(salary_band))
  )
  
salary_plot <- ggplot(salary_chart_df,
                       aes(x = chart_label, y = value, fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Reported" = unname(govukhugo::govuk_colours["blue"]),
      "Not reported" = unname(govukhugo::govuk_colours["dark-grey"])
    )
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1, margin = margin(r = 3)),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  salary_plot,
  width = 800,
  height = 450,
  alt_title = "Civil Service headcount by salary band",
  alt_desc = paste(
    "A bar chart showing the headcount of the Civil Service by salary band.",
    "The first bar is for those earning £150,000 or more per year, each",
    "subsequent bar shows the number of staff earning in bands of £5,000 until",
    "a bar for those earing £20,000 or less per year. The final bar shows the number",
    "whose salary was not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="salary-table">}}
```{r salary_table}

salary_band_table <- salary_band_all %>%
  select(salary_band, value, pc)

govukhugo::govuk_datatable(
    salary_band_table,
    title = "Civil Service headcount by salary band",
    col_names = c("Salary band", "Headcount", "Percent of known salary"),
    page_length = 10,
    search = FALSE
  ) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)

```
{{< /tab >}}
{{< tab title="High earners" heading="hide" ref="high-earners-table">}}
```{r high_earners_table}

high_earners %>%
  mutate(value = scales::comma(value, 1)) %>%
  knitr::kable(
    output = "html",
    col.names = c("Salary band", "Headcount"),
    align = "lr",
    caption = "Civil Service headcount in high earning salary bands"
  )

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

The median salary for all civil servants is `r salary_stats_out["median"]`, while the mean salary is `r salary_stats_out["mean"]`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="salary-stats-chart">}}
```{r salary_stats_chart}

salary_stats_df <- acses2021 %>%
  filter(source_table == "table_06a",
         sex == "All employees") %>%
  select(working_pattern, value_type, value) %>%
  mutate(
    working_pattern = factor(working_pattern, 
                             levels = c("Part-time", "Full-time", "All employees")),
    label = paste(
      if_else(
        str_detect(value_type, "_"),
        toupper(str_replace(value_type, "^(\\w{1})\\w+_(\\w{1})\\w+$", "\\1\\2")),
        str_to_title(value_type)
      ),
      gbp_format(value), sep = ": "
    ),
    label = str_wrap(label, 5),
    label_adjust = if_else(value_type == "lower_quartile" | value_type == "mean",
                           as.numeric(working_pattern) + 0.25,
                           as.numeric(working_pattern) - 0.25)
  )

salary_lines <- salary_stats_df %>%
  filter(value_type == "lower_quartile" | value_type == "upper_quartile") %>%
  select(working_pattern, value_type, value) %>%
  pivot_wider(names_from = value_type, values_from = value)

salary_points <- salary_stats_df %>%
  filter(value_type == "mean" | value_type == "median")
  
salary_stats_plot <- ggplot(salary_stats_df, 
                            aes(x = working_pattern)) +
  geom_linerange(aes(ymin = 20000, ymax = 50000),
                 size = 0.25,
                 colour = govukhugo::govuk_colours["dark-grey"]) +
  geom_linerange(data = salary_lines,
                aes(ymin = lower_quartile, ymax = upper_quartile),
                size = 8, 
                colour = govukhugo::govuk_colours["blue"]) +
  geom_point(data = salary_points,
             aes(y = value, shape = value_type),
             colour = govukhugo::govuk_colours["white"],
             size = 4) +
  geom_text(aes(label = label, 
                y = value,
                x = label_adjust),
            size = 3.5) +
  scale_shape_manual(
    values = c("median" = "circle",
               "mean" = "square")
  ) +
  scale_y_continuous(
    breaks = seq(20000, 50000, 5000),
    labels = c(gbp_format(20000), rep("", 5), gbp_format(50000))
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1, margin = margin(r = 3)),
    axis.ticks.x = element_line(
      colour = govukhugo::govuk_colours["govuk-border-colour"]
      ),
    axis.ticks.length.x = unit(4, "pt"),
    axis.text.x = element_text(
      size = 10, hjust = 0.5, margin = margin(r = 3),
      colour = govukhugo::govuk_colours["govuk-secondary-text-colour"]),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "none"
  )

govukhugo::render_svg(
  salary_stats_plot,
  width = 800,
  height = 400,
  alt_title = "Salary statistics by working pattern",
  alt_desc = paste(
    "A plot showing summary statistics of civil servant salaries by working",
    "pattern. Three rows show the statistics for all employees, full-time",
    "employees and part-time employees. Each row has a thin grey horizontal", 
    "line representing a range from £20,000 to £50,000, a thick blue horizontal",
    "line is overlaid showing the range from the lower quartile to upper",
    "quartile of salaries. Finally, two points are added to represent the median",
    "salary (a circle) and mean salary (a square)."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="stats-table">}}
```{r stats_table}

salary_stats_table <- salary_stats_df %>%
  select(working_pattern, value_type, value) %>%
  pivot_wider(names_from = value_type, values_from = value) %>%
  select(working_pattern, lower_quartile, median, mean, upper_quartile) %>%
  arrange(desc(working_pattern))

govukhugo::govuk_datatable(
    salary_stats_table,
    title = "Salary statistics by working pattern",
    col_names = c("Working pattern", "Lower quartile salary", 
                  "Median salary", "Mean salary", "Upper quartile salary"),
    page_length = 12,
    search = FALSE
  ) %>%
  DT::formatCurrency(columns = 2:5, currency = "£", digits = 0)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::
