---
title: "Data explorer"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: data
weight: 200
summary: "Data explorer main page"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

explorer_data <- readr::read_rds(file.path(data_dir, "acses_explorer.rds"))

page_vars <- c(
  "status", "year", "headcount", "fte", "mean_salary", "median_salary"
)

other_vars <- sort(names(explorer_data)[!(names(explorer_data) %in% page_vars)])

page_data <- explorer_data %>%
  filter(
    if_all(all_of(other_vars), ~.x == "Total"),
  ) %>%
  select(all_of(page_vars)) %>%
  mutate(
    across(c(headcount, fte),
           ~if_else(is.na(.x), "[c]", scales::comma(.x, accuracy = 1))),
    across(c(mean_salary, median_salary),
           ~if_else(is.na(.x), "[c]", scales::dollar(.x, accuracy = 1, prefix = "£")))
  )

var_labels <- c(
  "all" = "All civil servants",
  "status" = "Employment status",
  "year" = "Year",
  "age_band" = "Age band",
  "dept_group" = "Departmental group",
  "disability" = "Disability",
  "ethnicity" = "Ethnicity",
  "function_of_post" = "Function of post",
  "profession_of_post" = "Profession of post",
  "responsibility_level" = "Responsibilty level",
  "sex" = "Sex",
  "sexual_orientation" = "Sexual orientation",
  "uk_region" = "UK country or region",
  "headcount" = "Headcount",
  "fte" = "FTE",
  "mean_salary" = "Mean salary",
  "median_salary" = "Median salary"
)

breadcrumb <- function(var, depth, var_labels) {
  
  if (depth == 0) {
    depth_href = "#"
  } else {
    depth_href = paste0(rep("../", depth), collapse = "")
  }
  
  htmltools::tags$li(
    class = "govuk-breadcrumbs__list-item",
    htmltools::tags$a(
      class = "govuk-breadcrumbs__link",
      href = depth_href,
      var_labels[var]
    )
  )
}

further_link <- function(var, var_labels) {
  as.character(htmltools::tags$a(
      class = "govuk-link govuk-link--no-visited-state",
      href = paste0(var, "/"),
      var_labels[var])
  )
}

```

The data explorer provides access a range of further breakdowns of the underlying raw data that makes up the Civil Service Statistics. The explorer provides tables showing the headcount, full-time equivalent (FTE), mean and median salary for groups of civil servants, you are able to progressively select up to three variables to produce a breakdown of your results. The explorer tables show the figures for 2020 and 2021, and for staff in post, entrants and leavers.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-one-quarter .data-filters}
```{r}
htmltools::tagList(
  htmltools::tags$div(
    class = "govuk-breadcrumbs",
    htmltools::tags$ol(
      class = "govuk-breadcrumbs__list",
      breadcrumb(var = "all", depth = 0, var_labels = var_labels)
    )
  )
)

htmltools::tagList(
  htmltools::tags$div(
    class = "govuk-body-s",
    "Futher breakdowns available: ",
    glue::glue_collapse(
      purrr::map_chr(
        other_vars, ~further_link(.x, var_labels)
      ),
      sep = ", ", last = " or "
    )
  )
)
```
:::
:::{.govuk-grid-column-three-quarters .data-display}
```{r}
page_col_names <- unname(var_labels[names(page_data)])

govukhugo::govuk_datatable(
  page_data,
  col_names = page_col_names,
  small_text = TRUE,
  page_length = 10,
  col_defs = list(
    list(
      className = "dt-right",
      targets = (ncol(page_data) - 4):(ncol(page_data) - 1)
    )
  )
)
```
:::
::::

{{< break type="m" visible=true >}}
<h2 class="govuk-heading-s">Notes</h2>
{{< small >}}
Figures are rounded to the nearest 10 for headcount and full-time equivalent (FTE) and the nearest £100 for mean and median salary. Figures are suppressed when information relates to less than 50 civil servants (shown by the shorthand notiation "[c]").
{{< /small >}}
