---
title: "Civil servants by disability status"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: diversity
weight: 400
summary: "Headline statistics for civil servants by disability status"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

disability_grade <- acses2021 %>%
  filter(source_table == "table_03" & 
           disability != "All employees" &
           disability != "All declared") %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants", "Not reported",
      "All employees"
    ),
    disability = fct_relevel(
      disability, "Disabled", "Non-disabled", "Not declared", 
      "Not reported"
    ),
    valid = if_else(
      disability == "Not declared" | disability == "Not reported",
      NA_real_,
      value
    )
  ) %>%
  group_by(responsibility_level) %>%
  mutate(
    pc_disabled = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  select(responsibility_level, disability, value, pc_disabled) %>%
  arrange(responsibility_level, disability)

disability_totals <- disability_grade %>%
  filter(responsibility_level == "All employees") %>%
  select(-responsibility_level)

```

As of 31 March 2021, there were `r scales::comma(disability_totals$value[disability_totals$disability == "Disabled"], 1)` civil servants that reported they were disabled (equating to `r scales::percent(disability_totals$pc_disabled[disability_totals$disability == "Disabled"], 0.1)` of those whose disability status is known).[^1]

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="disability-chart">}}
```{r disability_chart}
disability_plot_df <- disability_totals %>%
  mutate(
    chart_label = paste0(disability, " (", scales::comma(value, 1), ")"),
    chart_label = fct_reorder(chart_label, as.numeric(disability)),
    fill_group = if_else(disability == "Not declared" | disability == "Not reported",
                         "Not provided",
                         "Provided")
  )

disability_plot <- ggplot(disability_plot_df, aes(x = fct_rev(chart_label), y = value, 
                                                fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Provided" = unname(govukhugo::govuk_colours["light-purple"]),
      "Not provided" = unname(govukhugo::govuk_colours["mid-grey"])
    )) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  disability_plot,
  width = 800,
  height = 100,
  alt_title = "Civil Service headcount by disability status",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by disability status",
    "After the bars for ethnicity there",
    "is one bar to show those whose have chosen not to declare their ethnicity",
    "and one bar to show those whose ethnicity is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="disability-table">}}
```{r disability_table}
govukhugo::govuk_datatable(
  disability_totals,
  title = "Civil Service headcount by disability status",
  col_names = c("Disability status", "Headcount", "Percent"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

At the Executive Officer responsibility level `r scales::percent(disability_grade$pc_disabled[disability_grade$responsibility_level == "Executive Officers" & disability_grade$disability == "Disabled"], 0.1)` have declared they are disabled, almost double the proportion of those at Senior Civil Service Level that declared they are disabled (`r scales::percent(disability_grade$pc_disabled[disability_grade$responsibility_level == "Senior Civil Service Level" & disability_grade$disability == "Disabled"], 0.1)`).

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="disability-grade-chart">}}
```{r disability_grade_chart}

disability_grade_plot_df <- disability_grade %>%
  filter(responsibility_level != "All employees" &
           responsibility_level != "Not reported") %>%
  drop_na(pc_disabled) %>%
  mutate(pc_label = scales::percent(pc_disabled, 1))

disability_grade_plot <- ggplot(
  disability_grade_plot_df,
  aes(x = fct_rev(responsibility_level), y = pc_disabled, fill = disability)
  ) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "Non-disabled" = unname(govukhugo::govuk_colours["light-purple"]),
      "Disabled" = scales::gradient_n_pal(govukhugo::govuk_palette("light-purple"))(c(0.5))
    ),
    name = "Disability status"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  disability_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by disability status and responsibility level",
  alt_desc = paste(
    "A stacked bar chart showing Civil Service headcount by responsibility level",
    "and disability status Bars represent the proportion of civil servants in",
    "each responsibility level that have declared they are non-disabled",
    "(the bar starting on the left) and those that have declared they are",
    "disabled (the bar starting on the right). The bars are ordered",
    "by responsibility level with the most senior level at the top and",
    "administrative staff at the bottom. The chart does not include",
    "data for where either responsbility level or disability status are not",
    "declared or reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="disability-grade-table">}}
```{r disability_grade_table}

disability_grade_table <- disability_grade %>%
  filter(responsibility_level != "All employees")

govukhugo::govuk_datatable(
  disability_grade_table,
  title = "Civil Service headcount by responsibility level and disability status",
  col_names = c("Responsibility level", "Disability status", "Headcount", "Percent"),
    page_length = 8,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

```{r disability_dept_setup}

disability_org <- acses2021 %>%
  filter(source_table == "table_38" &
           organisation != "All employees" &
           disability != "All employees" &
           disability != "All declared") %>%
  left_join(group_lookup, by = "organisation") %>%
  select(dept_group, organisation, disability, value) %>%
  mutate(
    disability = fct_relevel(
      disability, "Disabled", "Non-disabled", "Not declared", 
      "Not reported"
    )
  )

disability_dept <- disability_org %>%
  group_by(dept_group, disability) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(
    valid = if_else(disability == "Disabled" | disability == "Non-disabled",
                    value, 
                    NA_real_),
    pc_disability = valid/sum(valid, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  add_count(dept_group, wt = value, name = "group_reported")

disability_top3 <- disability_dept %>%
  filter(disability == "Disabled" & group_reported > 1500) %>%
  slice_max(pc_disability, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_disability, 3), 0.1),
      ")"
    )
  )

disability_bottom3 <- disability_dept %>%
  filter(disability == "Disabled" & group_reported > 1500) %>%
  slice_min(pc_disability, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_disability, 3), 0.1),
      ")"
    )
  )

```

The three departmental groups with the largest proportion of staff reporting they have a disability are: `r glue::glue_collapse(disability_top3$out_text, sep = "; ", last = " and, ")`. The three departmental groups with the smallest proportion of staff reporting they have a disability are: `r glue::glue_collapse(disability_bottom3$out_text, sep = "; ", last = " and, ")`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="disabilty-dept-chart">}}
```{r disability_dept_chart}

disability_dept_plot_df <- disability_dept %>%
  filter(disability == "Disabled" | disability == "Non-disabled") %>%
  mutate(
    dept_chart_group = case_when(
      group_reported < 1500 ~ "Other departments and agencies",
      grepl("Standards in Education", dept_group) ~ "Ofsted",
      TRUE ~ dept_group),
    dept_chart_group = fct_reorder(dept_chart_group, group_reported, .fun = max)
  ) %>%
  group_by(dept_chart_group, disability) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_disability = value/sum(value)) %>%
  ungroup() %>%
  mutate(
    pc_label = if_else(
      pc_disability < 0.1,
      NA_character_,
      scales::percent(janitor::round_half_up(pc_disability, 2), 1)
    )
  )

disability_dept_plot <- ggplot(
  disability_dept_plot_df,
  aes(x = dept_chart_group, y = pc_disability, fill = disability)
  ) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "Non-disabled" = unname(govukhugo::govuk_colours["light-purple"]),
      "Disabled" = scales::gradient_n_pal(govukhugo::govuk_palette("light-purple"))(c(0.5))
    ),
    name = "Disability status"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  disability_dept_plot,
  width = 800,
  height = 600,
  alt_title = "Civil Service headcount by disability status and departmental group",
  alt_desc = paste(
    "TO WRITE"
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Departmental groups" heading="hide" ref="disabilty-dept-table">}}
```{r disability_dept_table}
disability_dept_table <- disability_dept %>%
  select(dept_group, disability, value, pc_disability) %>%
  arrange(dept_group, disability)

govukhugo::govuk_datatable(
  disability_dept_table,
  title = "Civil Service headcount by departmental group and disability status",
  col_names = c("Departmental group", "Disability status", "Headcount", 
                "Percent of departmental group"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)
```
{{< /tab >}}
{{< tab title="Organisations" heading="hide" ref="disabilty-org-table">}}
```{r disability_org_table}

disability_org_table <- disability_org %>%
  mutate(valid = if_else(disability == "Disabled" | disability == "Non-disabled",
                         value,
                         NA_real_)
  ) %>%
  group_by(organisation) %>%
  mutate(
    pc_disability = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  select(dept_group, organisation, disability, value, pc_disability) %>%
  arrange(dept_group, organisation, disability)

govukhugo::govuk_datatable(
  disability_org_table,
  title = "Civil Service headcount by organisation and disability status",
  col_names = c("Departmental group", "Organisation", "Disability status",
                "Headcount", "Percent of organisation"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 4, digits = 0) %>%
  DT::formatPercentage(columns = 5, digits = 1)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

[^1]: The Civil Service Statistics record two types of missing value for disability status. Where an individual has specifically chosen not to provide their disability status, for example by selecting a "Prefer not to say" value, this is recorded as "Not declared". Where an individual has not provided any data on their disability status, including not selecting a "Prefer not to say" value, this is recorded as "Not reported".
