---
title: "Civil servants by ethnicity and national identity"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: diversity
weight: 300
summary: "Headline statistics for civil servants by ethnicity and national identity"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

# ACSES 2021 separates Chinese from Asian, this is not in line with GSS
# harmonised standards
ethnicity_corrected <- acses2021 %>%
  filter(source_table == "table_02") %>%
  select(responsibility_level, ethnicity, value) %>%
  mutate(ethnicity = if_else(ethnicity == "Chinese", "Asian", ethnicity)) %>%
  group_by(responsibility_level, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop")

ethnicity_totals <- ethnicity_corrected %>%
  filter(
      responsibility_level == "All employees" &
      ethnicity %in% c("White", "Asian", "Black", "Mixed", "Other")) %>%
  group_by(white = ethnicity == "White") %>%
  summarise(value = sum(value), .groups = "drop") %>%
  mutate(
    pc_eth = value/sum(value),
    out_num = scales::comma(value),
    out_pc = scales::percent(janitor::round_half_up(pc_eth, 3), 0.1)
  )

```

As at 31 March 2021 there were `r ethnicity_totals$out_num[ethnicity_totals$white == TRUE]` civil servants that reported have a white ethnicity (equating to `r ethnicity_totals$out_pc[ethnicity_totals$white == TRUE]` of those with a known ethnicity) and there were `r ethnicity_totals$out_num[ethnicity_totals$white == FALSE]` civil servants that reported the were from an ethnic minority background (equating to `r ethnicity_totals$out_pc[ethnicity_totals$white == FALSE]` of those with a known ethnicity).

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-chart">}}
```{r ethnicity_chart}
ethnicity_plot_df <-  ethnicity_corrected %>%
  filter(
      responsibility_level == "All employees" &
      ethnicity != "All employees" &
      ethnicity != "All declared"
  ) %>%
  mutate(
    ethnicity = fct_relevel(ethnicity, "White", "Mixed", "Asian", "Black",
                            "Other", "Not declared", "Not reported"),
    fill_group = if_else(
      ethnicity == "Not declared" | ethnicity == "Not reported",
      "Not provided",
      "Provided"
    ),
    chart_label = paste0(ethnicity, " (", scales::comma(value), ")"),
    chart_label = fct_rev(fct_reorder(chart_label, as.numeric(ethnicity)))
  ) %>%
  arrange(chart_label)

ethnicity_plot <- ggplot(ethnicity_plot_df, aes(x = chart_label, y = value, 
                                                fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Provided" = unname(govukhugo::govuk_colours["red"]),
      "Not provided" = unname(govukhugo::govuk_colours["mid-grey"])
    )) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  ethnicity_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by ethnicity",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by ethnicity.",
    "The bars are ordered in accordance with the Government Statistical",
    "Service's Harmonised Standard for ethnicity information, the first",
    "category is for those with a White ethnicity followed by Mixed, Asian,",
    "Black and finally Other ethnicities. After the bars for ethnicity there",
    "is one bar to show those whose have chosen not to declare their ethnicity",
    "and one bar to show those whose ethnicity is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="ethnicity-table">}}
```{r ethnicity_table}
ethnicity_table <- ethnicity_plot_df %>%
  select(ethnicity, value) %>%
  mutate(
    valid = if_else(
      ethnicity %in% c("White", "Mixed", "Asian", "Black", "Other"),
      value,
      NA_real_
    ),
    pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3),
  ) %>%
  arrange(ethnicity) %>%
  select(-valid)

govukhugo::govuk_datatable(
  ethnicity_table,
  title = "Civil Service headcount by ethnicity",
  col_names = c("Ethnicity", "Headcount", "Percent"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

Around one in nine of those at Senior Civil Service Level are from an ethnic 
minority background, compared to around one in six of those at the Executive 
Officer responsibility level.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-grade-chart">}}
```{r ethnicity_grade_chart}
ethnicity_grade_plot_df <- ethnicity_corrected %>%
  filter((ethnicity %in% c("White", "Asian", "Black", "Mixed", "Other")) &
           responsibility_level != "All employees" &
           responsibility_level != "Not reported") %>%
  mutate(
    ethnicity = if_else(ethnicity == "White", "White", "Ethnic minority"),
    responsibility_level = fct_rev(factor(
      responsibility_level,
      levels = c("Senior Civil Service Level", "Grades 6 and 7", 
                 "Senior and Higher Executive Officers", "Executive Officers",
                 "Administrative Officers and Assistants")))
  ) %>%
  group_by(responsibility_level, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  mutate(pc_label = scales::percent(janitor::round_half_up(pc_eth, 2), 1))

ethnicity_grade_plot <- ggplot(ethnicity_grade_plot_df,
                               aes(x = responsibility_level, y = value,
                                   fill = ethnicity)) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(
    values = c(
      "White" = unname(govukhugo::govuk_colours["red"]),
      "Ethnic minority" = scales::gradient_n_pal(govukhugo::govuk_palette("red"))(c(0.5))
    ),
    name = "Ethnicity"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  ethnicity_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by ethnicity and responsibility level",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by ethnicity."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="ethnicity-grade-table">}}
```{r ethnicity_grade_table}
ethnicity_grade_table <- ethnicity_corrected %>%
  filter(ethnicity != "All declared" & ethnicity != "All employees" &
           responsibility_level != "All employees") %>%
  mutate(
    ethnicity = fct_relevel(
      ethnicity, 
      "White", "Mixed", "Asian", "Black", "Other", "Not declared", "Not reported"
    ),
    responsibility_level = factor(
      responsibility_level,
      levels = c("Senior Civil Service Level", "Grades 6 and 7", 
                 "Senior and Higher Executive Officers", "Executive Officers",
                 "Administrative Officers and Assistants", "Not reported")
    ),
    valid = if_else(ethnicity == "Not declared" | ethnicity == "Not reported",
                    NA_real_,
                    value)
  ) %>%
  group_by(responsibility_level) %>%
  mutate(pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)) %>%
  select(responsibility_level, ethnicity, value, pc_eth) %>%
  arrange(responsibility_level, ethnicity)

govukhugo::govuk_datatable(
  ethnicity_grade_table,
  title = "Civil Service headcount by responsibility level and ethnicity",
  col_names = c("Responsibility level", "Ethnicity", "Headcount", 
                "Percent of responsibility level"),
    page_length = 7,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

```{r ethnicity_dept_setup}
ethnicity_org <- acses2021 %>%
  filter(source_table == "table_37" & organisation != "All employees") %>%
  left_join(group_lookup, by = "organisation") %>%
  select(dept_group, organisation, ethnicity, value)

ethnicity_dept <- ethnicity_org %>%
  filter(ethnicity %in% c("White", "Ethnic minority")) %>%
  group_by(dept_group, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  add_count(dept_group, wt = value, name = "group_reported")

ethnicity_dept_top3em <- ethnicity_dept %>%
  filter(ethnicity == "Ethnic minority" & group_reported > 1500) %>%
  slice_max(pc_eth, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_eth, 3), 0.1),
      ")"
    )
  )

ethnicity_dept_bottom3em <- ethnicity_dept %>%
  filter(ethnicity == "Ethnic minority" & group_reported > 1500) %>%
  slice_min(pc_eth, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_eth, 3), 0.1),
      ")"
    )
  )

```

The three departmental groups with the largest proportion of staff from an ethnic minority background are: `r glue::glue_collapse(ethnicity_dept_top3em$out_text, sep = "; ", last = " and, ")`. The three departmental groups with the smallest proportion of staff from an ethnic minority background are: `r glue::glue_collapse(ethnicity_dept_bottom3em$out_text, sep = "; ", last = " and, ")`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-dept-chart">}}
```{r ethnicity_dept_chart}
ethnicity_dept_plot_df <- ethnicity_dept %>%
  mutate(
    dept_chart_group = case_when(
      group_reported < 1500 ~ "Other departments and agencies",
      grepl("Standards in Education", dept_group) ~ "Ofsted",
      TRUE ~ dept_group),
    dept_chart_group = fct_reorder(dept_chart_group, group_reported, .fun = max),
    ethnicity = fct_rev(fct_relevel(ethnicity, "White", "Ethnic minority"))
  ) %>%
  group_by(dept_chart_group, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  mutate(
    chart_label = if_else(
      pc_eth < 0.1,
      NA_character_,
      scales::percent(janitor::round_half_up(pc_eth, 2), 1)
    )
  )

ethnicity_dept_plot <- ggplot(ethnicity_dept_plot_df,
                              aes(x = dept_chart_group, y = pc_eth,
                                  fill = ethnicity)) +
  geom_col(position = "fill") +
  geom_text(aes(label = chart_label), 
            position = position_fill(vjust = 0.5),
            show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "White" = unname(govukhugo::govuk_colours["red"]),
      "Ethnic minority" = scales::gradient_n_pal(govukhugo::govuk_palette("red"))(c(0.5))
    ),
    name = "Ethnicity"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  ethnicity_dept_plot,
  width = 800,
  height = 600,
  alt_title = "Civil Service headcount by ethnicity and departmental group",
  alt_desc = paste(
    "TO WRITE"
  ),
  caption = "alt_title"
)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::
