---
title: "Civil servants by ethnicity and national identity"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: diversity
weight: 300
summary: "Headline statistics for civil servants by ethnicity and national identity"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

```

This page provides statistics on civil servants by [ethnicity](#ethnicity) and [national identity](#national-identity)

## Ethnicity

```{r ethnicity_setup}
# ACSES 2021 separates Chinese from Asian, this is not in line with GSS
# harmonised standards
ethnicity_corrected <- acses2021 %>%
  filter(source_table == "table_02") %>%
  select(responsibility_level, ethnicity, value) %>%
  mutate(ethnicity = if_else(ethnicity == "Chinese", "Asian", ethnicity)) %>%
  group_by(responsibility_level, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop")

ethnicity_totals <- ethnicity_corrected %>%
  filter(
      responsibility_level == "All employees" &
      ethnicity %in% c("White", "Asian", "Black", "Mixed", "Other")) %>%
  group_by(white = ethnicity == "White") %>%
  summarise(value = sum(value), .groups = "drop") %>%
  mutate(
    pc_eth = value/sum(value),
    out_num = scales::comma(value),
    out_pc = scales::percent(janitor::round_half_up(pc_eth, 3), 0.1)
  )
```

As at 31 March 2021 there were `r ethnicity_totals$out_num[ethnicity_totals$white == TRUE]` civil servants that reported have a white ethnicity (equating to `r ethnicity_totals$out_pc[ethnicity_totals$white == TRUE]` of those with a known ethnicity) and there were `r ethnicity_totals$out_num[ethnicity_totals$white == FALSE]` civil servants that reported the were from an ethnic minority background (equating to `r ethnicity_totals$out_pc[ethnicity_totals$white == FALSE]` of those with a known ethnicity).[^1]

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-chart">}}
```{r ethnicity_chart}
ethnicity_plot_df <-  ethnicity_corrected %>%
  filter(
      responsibility_level == "All employees" &
      ethnicity != "All employees" &
      ethnicity != "All declared"
  ) %>%
  mutate(
    ethnicity = fct_relevel(ethnicity, "White", "Mixed", "Asian", "Black",
                            "Other", "Not declared", "Not reported"),
    fill_group = if_else(
      ethnicity == "Not declared" | ethnicity == "Not reported",
      "Not provided",
      "Provided"
    ),
    chart_label = paste0(ethnicity, " (", scales::comma(value), ")"),
    chart_label = fct_rev(fct_reorder(chart_label, as.numeric(ethnicity)))
  ) %>%
  arrange(chart_label)

ethnicity_plot <- ggplot(ethnicity_plot_df, aes(x = chart_label, y = value, 
                                                fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Provided" = unname(govukhugo::govuk_colours["bright-purple"]),
      "Not provided" = unname(govukhugo::govuk_colours["mid-grey"])
    )) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  ethnicity_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by ethnicity",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by ethnicity.",
    "The bars are ordered in accordance with the Government Statistical",
    "Service's Harmonised Standard for ethnicity information, the first",
    "category is for those with a White ethnicity followed by Mixed, Asian,",
    "Black and finally Other ethnicities. After the bars for ethnicity there",
    "is one bar to show those whose have chosen not to declare their ethnicity",
    "and one bar to show those whose ethnicity is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="ethnicity-table">}}
```{r ethnicity_table}
ethnicity_table <- ethnicity_plot_df %>%
  select(ethnicity, value) %>%
  mutate(
    valid = if_else(
      ethnicity %in% c("White", "Mixed", "Asian", "Black", "Other"),
      value,
      NA_real_
    ),
    pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3),
  ) %>%
  arrange(ethnicity) %>%
  select(-valid)

govukhugo::govuk_datatable(
  ethnicity_table,
  title = "Civil Service headcount by ethnicity",
  col_names = c("Ethnicity", "Headcount", "Percent"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

Around one in nine of those at Senior Civil Service Level are from an ethnic 
minority background, compared to around one in six of those at the Executive 
Officer responsibility level.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-grade-chart">}}
```{r ethnicity_grade_chart}
ethnicity_grade_plot_df <- ethnicity_corrected %>%
  filter((ethnicity %in% c("White", "Asian", "Black", "Mixed", "Other")) &
           responsibility_level != "All employees" &
           responsibility_level != "Not reported") %>%
  mutate(
    ethnicity = if_else(ethnicity == "White", "White", "Ethnic minority"),
    responsibility_level = fct_rev(factor(
      responsibility_level,
      levels = c("Senior Civil Service Level", "Grades 6 and 7", 
                 "Senior and Higher Executive Officers", "Executive Officers",
                 "Administrative Officers and Assistants")))
  ) %>%
  group_by(responsibility_level, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  mutate(pc_label = scales::percent(janitor::round_half_up(pc_eth, 2), 1))

ethnicity_grade_plot <- ggplot(ethnicity_grade_plot_df,
                               aes(x = responsibility_level, y = value,
                                   fill = ethnicity)) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "White" = unname(govukhugo::govuk_colours["bright-purple"]),
      "Ethnic minority" = scales::gradient_n_pal(govukhugo::govuk_palette("bright-purple"))(c(0.5))
    ),
    name = "Ethnicity"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  ethnicity_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by ethnicity and responsibility level",
  alt_desc = paste(
    "A stacked bar chart showing Civil Service headcount by responsibility level",
    "and ethnicity. Bars represent the proportion of civil servants in",
    "each responsibility level that have declared their ethnicity as White",
    "(the bar starting on the left) and those that have declared their ethnicity",
    "as an ethnic minority (the bar starting on the right). The bars are ordered",
    "by responsibility level with the most senior level at the top and",
    "administrative staff at the bottom. The chart does not include",
    "data for where either responsbility level or ethnicity are not",
    "declared or reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="ethnicity-grade-table">}}
```{r ethnicity_grade_table}
ethnicity_grade_table <- ethnicity_corrected %>%
  filter(ethnicity != "All declared" & ethnicity != "All employees" &
           responsibility_level != "All employees") %>%
  mutate(
    ethnicity = fct_relevel(
      ethnicity, 
      "White", "Mixed", "Asian", "Black", "Other", "Not declared", "Not reported"
    ),
    responsibility_level = factor(
      responsibility_level,
      levels = c("Senior Civil Service Level", "Grades 6 and 7", 
                 "Senior and Higher Executive Officers", "Executive Officers",
                 "Administrative Officers and Assistants", "Not reported")
    ),
    valid = if_else(ethnicity == "Not declared" | ethnicity == "Not reported",
                    NA_real_,
                    value)
  ) %>%
  group_by(responsibility_level) %>%
  mutate(pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)) %>%
  select(responsibility_level, ethnicity, value, pc_eth) %>%
  arrange(responsibility_level, ethnicity)

govukhugo::govuk_datatable(
  ethnicity_grade_table,
  title = "Civil Service headcount by responsibility level and ethnicity",
  col_names = c("Responsibility level", "Ethnicity", "Headcount", 
                "Percent of responsibility level"),
    page_length = 7,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

```{r ethnicity_dept_setup}
ethnicity_org <- acses2021 %>%
  filter(source_table == "table_37" & organisation != "All employees") %>%
  left_join(group_lookup, by = "organisation") %>%
  select(dept_group, organisation, ethnicity, value)

ethnicity_dept <- ethnicity_org %>%
  filter(ethnicity %in% c("White", "Ethnic minority")) %>%
  group_by(dept_group, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  add_count(dept_group, wt = value, name = "group_reported")

ethnicity_dept_top3em <- ethnicity_dept %>%
  filter(ethnicity == "Ethnic minority" & group_reported > 1500) %>%
  slice_max(pc_eth, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_eth, 3), 0.1),
      ")"
    )
  )

ethnicity_dept_bottom3em <- ethnicity_dept %>%
  filter(ethnicity == "Ethnic minority" & group_reported > 1500) %>%
  slice_min(pc_eth, n = 3) %>%
  mutate(
    out_text = paste0(
      dept_group, " (", 
      scales::percent(janitor::round_half_up(pc_eth, 3), 0.1),
      ")"
    )
  )

```

The three departmental groups with the largest proportion of staff from an ethnic minority background are: `r glue::glue_collapse(ethnicity_dept_top3em$out_text, sep = "; ", last = " and, ")`. The three departmental groups with the smallest proportion of staff from an ethnic minority background are: `r glue::glue_collapse(ethnicity_dept_bottom3em$out_text, sep = "; ", last = " and, ")`.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="ethnicity-dept-chart">}}
```{r ethnicity_dept_chart}
ethnicity_dept_plot_df <- ethnicity_dept %>%
  mutate(
    dept_chart_group = case_when(
      group_reported < 1500 ~ "Other departments and agencies",
      grepl("Standards in Education", dept_group) ~ "Ofsted",
      TRUE ~ dept_group),
    dept_chart_group = fct_reorder(dept_chart_group, group_reported, .fun = max),
    ethnicity = fct_rev(fct_relevel(ethnicity, "White", "Ethnic minority"))
  ) %>%
  group_by(dept_chart_group, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_eth = value/sum(value)) %>%
  ungroup() %>%
  mutate(
    chart_label = if_else(
      pc_eth < 0.1,
      NA_character_,
      scales::percent(janitor::round_half_up(pc_eth, 2), 1)
    )
  )

ethnicity_dept_plot <- ggplot(ethnicity_dept_plot_df,
                              aes(x = dept_chart_group, y = pc_eth,
                                  fill = ethnicity)) +
  geom_col(position = "fill") +
  geom_text(aes(label = chart_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "White" = unname(govukhugo::govuk_colours["bright-purple"]),
      "Ethnic minority" = scales::gradient_n_pal(govukhugo::govuk_palette("bright-purple"))(c(0.5))
    ),
    name = "Ethnicity"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  ethnicity_dept_plot,
  width = 800,
  height = 600,
  alt_title = "Civil Service headcount by ethnicity and departmental group",
  alt_desc = paste(
    "TO WRITE"
  ),
  caption = "alt_title"
)
```
{{< /tab >}}
{{< tab title="Departmental groups" heading="hide" ref="ethnicity-dept-table">}}
```{r ethnicity_dept_table}
ethnicity_dept_table <- ethnicity_org %>%
  filter(ethnicity != "All declared" & ethnicity != "All employees") %>%
  mutate(ethnicity = fct_relevel(ethnicity, "White", "Ethnic minority",
                                 "Not declared", "Not reported")) %>%
  group_by(dept_group, ethnicity) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(
    valid = if_else(
      ethnicity == "White" | ethnicity == "Ethnic minority",
      value,
      NA_real_),
    pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  select(-valid) %>%
  arrange(dept_group, ethnicity)

govukhugo::govuk_datatable(
  ethnicity_dept_table,
  title = "Civil Service headcount by departmental group and ethnicity",
  col_names = c("Departmental group", "Ethnicity", "Headcount", 
                "Percent of departmental group"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)
```
{{< /tab >}}
{{< tab title="Organisations" heading="hide" ref="ethnicity-org-table">}}
```{r ethnicity_org_table}
ethnicity_org_table <- ethnicity_org %>%
  filter(ethnicity != "All declared" & ethnicity != "All employees") %>%
  mutate(ethnicity = fct_relevel(ethnicity, "White", "Ethnic minority",
                                 "Not declared", "Not reported")) %>%
  group_by(organisation) %>%
  mutate(
    valid = if_else(
      ethnicity == "White" | ethnicity == "Ethnic minority",
      value,
      NA_real_),
    pc_eth = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  select(-valid) %>%
  arrange(dept_group, organisation, ethnicity)

govukhugo::govuk_datatable(
  ethnicity_org_table,
  title = "Civil Service headcount by organisation and ethnicity",
  col_names = c("Departmental group", "Organisation", "Ethnicity", "Headcount", 
                "Percent of organisation"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 4, digits = 0) %>%
  DT::formatPercentage(columns = 5, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

## National identity

```{r natid_setup}
# ACSES 2021 separates Chinese from Asian, this is not in line with GSS
# harmonised standards
natid_grade <- acses2021 %>%
  filter(source_table == "table_05" & 
           national_identity != "All employees") %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants", "Not reported",
      "All employees"
    ),
    national_identity = fct_relevel(
      national_identity, "British or Mixed British", "English", "Northern Irish",
      "Scottish", "Welsh", "Other national identity", "Not declared", 
      "Not reported"
    ),
    valid = if_else(
      national_identity == "Not declared" | national_identity == "Not reported",
      NA_real_,
      value
    )
  ) %>%
  group_by(responsibility_level) %>%
  mutate(
    pc_natid = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  select(responsibility_level, national_identity, value, pc_natid) %>%
  arrange(responsibility_level, national_identity)

natid_totals <- natid_grade %>%
  filter(responsibility_level == "All employees") %>%
  select(-responsibility_level)

natid_missing <- natid_grade %>%
  mutate(
    status = if_else((national_identity == "Not declared" | 
                       national_identity == "Not reported"),
                     "Not provided",
                     "Provided")
  ) %>%
  group_by(responsibility_level, status) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_miss = value/sum(value)) %>%
  ungroup() %>%
  arrange(responsibility_level, status)
  

```

As at 31 March 2021 there were `r scales::comma(natid_totals$value[natid_totals$national_identity == "Other national identity"], 1)` civil servants that had a national identity from outside the United Kingdom[^2], equating to `r scales::percent(natid_totals$pc_natid[natid_totals$national_identity == "Other national identity"], 0.1)` of those with a known national identity.[^3]

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="natid-chart">}}
```{r natid_chart}

natid_plot_df <- natid_totals %>%
  mutate(
    chart_label = paste0(national_identity, " (", scales::comma(value, 1), ")"),
    chart_label = fct_rev(fct_reorder(chart_label, as.numeric(national_identity)))
  )

natid_plot_colours <- c(
  scales::gradient_n_pal(govukhugo::govuk_palette("blues"))(scales::rescale(1:5)),
  govukhugo::govuk_colours["orange"],
  rep(govukhugo::govuk_colours["mid-grey"], 2)
)

names(natid_plot_colours) <- levels(natid_totals$national_identity)

natid_plot <- ggplot(natid_plot_df,
                     aes(x = chart_label, y = value, fill = national_identity)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = natid_plot_colours) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  natid_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by national identity",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by national identity",
    "The bars are ordered with the bar for British or Mixed British first",
    "followed by bars for other UK national identities in alphabetical order",
    "(English, Northern Irish, Scottish, Welsh) followed by a bar for those",
    "with another national identity (that is, they do not have a UK national",
    "identity). Finally two bars to show missing values, one bar to show those",
    "whose have chosen not to declare their national identity and one bar to",
    "show those whose national identity is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="natid-table">}}
```{r natid_table}

govukhugo::govuk_datatable(
  natid_totals,
  title = "Civil Service headcount by national identity",
  col_names = c("National identity", "Headcount", "Percent"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)

```
{{< /tab>}}
{{< /tabset >}}
:::
::::

The proportion of staff with known national identity who declare a national identity from outside the United Kingdom is broadly similar across responsibility levels. However, at junior responsibility levels data on national identity is less likely to be available than for those at senior levels.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="natid-grade-chart">}}
```{r natid_grade_chart}

natid_grade_plot_df <- natid_grade %>%
  filter(responsibility_level != "Not reported" & 
           responsibility_level != "All employees" &
           national_identity != "British or Mixed British") %>%
  drop_na(pc_natid) %>%
  mutate(chart_label = if_else(pc_natid < 0.02, NA_character_, 
                               scales::percent(pc_natid, 1))) %>%
  arrange(responsibility_level, national_identity)

natid_grade_plot_colours <- c(
  scales::gradient_n_pal(govukhugo::govuk_palette("blues"))(scales::rescale(1:5)),
  govukhugo::govuk_colours["orange"]
)

natid_grade_plot_colours <- natid_plot_colours[2:6]

natid_grade_plot <- ggplot(natid_grade_plot_df,
                     aes(x = fct_rev(responsibility_level), y = pc_natid, fill = fct_rev(national_identity))) +
  geom_col() +
  geom_text(aes(label = chart_label), position = position_stack(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = natid_grade_plot_colours,
    name = "National identity",
    guide = guide_legend(nrow = 2, direction = "horizontal", byrow = TRUE)
  ) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top"
  )

govukhugo::render_svg(
  natid_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by responsibility level and national identity",
  alt_desc = paste(
    "A stacked bar chart showing Civil Service headcount by responsibility level",
    "and national identity. Bars represent the proportion of civil servants in",
    "each responsibility level that have a declared national identity that is",
    "not British or Mixed British (that is, either English, Northern Irish,)",
    "Welsh, Scottish or another national identity. Bars are sorted by",
    "responsibility level with the most senior level at the top and",
    "administrative staff at the bottom. The chart does not include",
    "data for where either responsbility level or national identity are not",
    "declared or reported."
  ),
  caption = "alt_title"
)
```
<p class="govuk-body-s">Chart excludes those who declared their national identity as British or Mixed British.</p>
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="natid-grade-table">}}
```{r natid_grade_table}

natid_grade_table <- natid_grade %>%
  filter(responsibility_level != "All employees")

govukhugo::govuk_datatable(
  natid_grade_table,
  title = "Civil Service headcount by responsibility level and national identity",
  col_names = c("Responsibility level", "National identity", "Headcount", "Percent"),
    page_length = 8,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)

```
{{< /tab>}}
:::
::::


[^1]: The Civil Service Statistics record two types of missing value for ethnicity and national identity. Where an individual has specifically chosen not to provide their ethnicity or national identity, for example by selecting a "Prefer not to say" value, this is recorded as "Not declared". Where an individual has not provided any data on their ethnicity or national identity, including not selecting a "Prefer not to say" value, this is recorded as "Not reported".
[^2]: That is they have declared that their ethnicity is not any of the following: British or Mixed British, English, Northern Irish, Scottish, or Welsh.
[^3]: National identity has two missing values, "not declared" and "not reported", see endnote [1](#fn1) for more details. There are `r scales::comma(natid_missing$value[natid_missing$responsibility_level == "All employees" & natid_missing$status == "Not provided"], 1)` individuals whose national identity is either not declared or not reported, which is `r scales::percent(natid_missing$pc_miss[natid_missing$responsibility_level == "All employees" & natid_missing$status == "Not provided"], 0.1)` of total headcount.
