---
title: "Civil servants by religion or belief"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: diversity
weight: 600
summary: "Headline statistics for civil servants by religion or belief"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

religion_grade <- acses2021 %>%
  filter(source_table == "table_a4" &
           religion_or_belief != "All employees" & 
           religion_or_belief != "All declared") %>%
  mutate(
    religion_or_belief = fct_relevel(
      religion_or_belief, "No religion", "Christian", "Buddhist", "Hindu",
      "Jewish", "Muslim", "Sikh", "Any other religion", "Not declared", 
      "Not reported"
    ),
    major_group = case_when(
      religion_or_belief == "Christian" ~ "Christian",
      religion_or_belief == "No religion" ~ "No religion",
      religion_or_belief == "Not declared" ~ "Not provided",
      religion_or_belief == "Not reported" ~ "Not provided",
      TRUE ~ "Other religion"
    )
  )

religion_totals <- religion_grade %>%
  filter(responsibility_level == "All employees") %>%
  group_by(major_group) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  mutate(
    valid = if_else(major_group == "Not provided", NA_real_, value),
    pc_all = janitor::round_half_up(value/sum(value), 3),
    pc_valid = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  )

```

As at 31 March 2021, there were `r scales::comma(religion_totals$value[religion_totals$major_group == "Christian"], 1)` civil servants who declared their religion or belief as Christian, this equates to `r scales::percent(religion_totals$pc_valid[religion_totals$major_group == "Christian"], 0.1)` of those who declared their religion or belief. There were`r scales::comma(religion_totals$value[religion_totals$major_group == "Other religion"], 1)`civil servants who declared having another religion or belief (`r scales::percent(religion_totals$pc_valid[religion_totals$major_group == "Other religion"], 0.1)` of those who declared their religion), and `r scales::comma(religion_totals$value[religion_totals$major_group == "Other religion"], 1)`civil servants who declared they have no religion (`r scales::percent(religion_totals$pc_valid[religion_totals$major_group == "Other religion"], 0.1)` of those who declared their religion).[^1]

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="religion-chart">}}
```{r religion_chart}
religion_plot_df <- religion_grade %>%
  filter(responsibility_level == "All employees") %>%
  select(religion_or_belief, value) %>%
  mutate(
    fill_group = if_else(
      religion_or_belief == "Not declared" | religion_or_belief == "Not reported",
      "Not provided",
      "Provided"
    ),
    chart_label = paste0(religion_or_belief, " (", scales::comma(value, 1), ")"),
    chart_label = fct_reorder(chart_label, as.numeric(religion_or_belief))
  )

religion_plot <- ggplot(religion_plot_df, 
                        aes(x = fct_rev(chart_label), y = value, fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Provided" = unname(govukhugo::govuk_colours["dark-blue"]),
      "Not provided" = unname(govukhugo::govuk_colours["mid-grey"])
    )) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  religion_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by religion or belief",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by religion or belief.",
    "After the bars for religions there is one bar to show those",
    "whose have chosen not to declare their religion or belief and one bar to",
    "show those whose religion or belief is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="religion-table">}}
```{r religion_table}

religion_table <- religion_grade %>%
  filter(responsibility_level == "All employees") %>%
  mutate(
    valid = if_else(major_group == "Not provided", NA_real_, value),
    pc_religion = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
    ) %>%
  select(religion_or_belief, value, pc_religion) %>%
  arrange(religion_or_belief)

govukhugo::govuk_datatable(
  religion_table,
  title = "Civil Service headcount by religion or belief",
  col_names = c("Religion or belief", "Headcount", "Percent"),
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

The proportion of staff at all grades who report they are a Christian is similar across all responsibility levels, staff at junior grades are slightly less likely to declare they have no religion and more likely to declare they have another religion than those at senior grades.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="religion-grade-chart">}}
```{r religion_grade_chart}
religion_grade_plot_df <- religion_grade %>%
  filter(responsibility_level != "All employees" &
           responsibility_level != "Not reported" &
           major_group != "Not provided") %>%
  group_by(responsibility_level, major_group) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_religion = value/sum(value)) %>%
  ungroup() %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants"
    ),
    major_group = fct_relevel(major_group, "No religion", "Christian", "Other religion"),
    pc_label = scales::percent(janitor::round_half_up(pc_religion, 2), 1)
  )

religion_grade_plot <- ggplot(
  religion_grade_plot_df,
  aes(x = fct_rev(responsibility_level), y = pc_religion, fill = fct_rev(major_group))
  ) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "No religion" = unname(govukhugo::govuk_colours["dark-blue"]),
      "Christian" =
        scales::gradient_n_pal(govukhugo::govuk_palette("dark-blue"))(c(0.25)),
      "Other religion" =
        scales::gradient_n_pal(govukhugo::govuk_palette("dark-blue"))(c(0.5))
    ),
    name = "Religion or belief"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  religion_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by religion or belief and responsibility level",
  alt_desc = paste(
    "A stacked bar chart showing Civil Service headcount by responsibility level",
    "and religion or belief. Bars represent the proportion of civil servants in",
    "each responsibility level that have declared they have no religion",
    "(the bar starting on the left), those that have declared they are Christian", 
    "(the middle bar), and those that have declared they have another religion.",
    "The bars are ordered by responsibility level with the most",
    "senior level at the top and administrative staff at the bottom. The chart",
    "does not include data for where either responsbility level or sexual",
    "orientation are not declared or reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="religion-grade-table-chart">}}
```{r religion_grade_table}
religion_grade_table <- religion_grade %>%
  filter(responsibility_level != "All employees") %>%
  group_by(responsibility_level) %>%
  mutate(
    valid = if_else(major_group == "Not provided", NA_real_, value),
    pc_religion = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants", "Not reported"
    )
  ) %>%
  select(responsibility_level, religion_or_belief, value, pc_religion) %>%
  arrange(responsibility_level, religion_or_belief)

govukhugo::govuk_datatable(
  religion_grade_table,
  title = "Civil Service headcount by responsibility level and religion or belief",
  col_names = c("Responsibility level", "Religion or belief", "Headcount", "Percent"),
    page_length = 9,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

[^1]: The Civil Service Statistics record two types of missing value for religion or belief. Where an individual has specifically chosen not to provide their religion or belief, for example by selecting a "Prefer not to say" value, this is recorded as "Not declared". Where an individual has not provided any data on their religion or belief, including not selecting a "Prefer not to say" value, this is recorded as "Not reported". There are `r scales::comma(religion_totals$value[religion_totals$major_group == "Not provided"], 1)` individuals whose religion or belief is either not declared or not reported, which is `r scales::percent(religion_totals$pc_all[religion_totals$major_group == "Not provided"], 0.1)` of total headcount.
