---
title: "Civil servants by sexual orientation"
knit: govukhugo::govukhugo_knit
date: 2021-01-01
section: diversity
weight: 500
summary: "Headline statistics for civil servants by sexual orientation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

orientation_grade <- acses2021 %>%
  filter(source_table == "table_a2" &
           sexual_orientation != "All employees" & 
           sexual_orientation != "All declared") %>%
  mutate(
    sexual_orientation = fct_relevel(
      sexual_orientation, "Heterosexual/Straight", "Gay/Lesbian",
      "Bisexual", "Other", "Not declared", "Not reported"
    ),
    valid = if_else(
      sexual_orientation == "Not declared" | sexual_orientation == "Not reported",
      NA_real_,
      value
    ),
    major_group = case_when(
      sexual_orientation == "Not declared" ~ "Not provided",
      sexual_orientation == "Not reported" ~ "Not provided",
      sexual_orientation == "Heterosexual/Straight" ~ "Heterosexual",
      TRUE ~ "LGBO"
    )
  )

orientation_totals <- orientation_grade %>%
  filter(responsibility_level == "All employees") %>%
  group_by(major_group) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  mutate(
    valid = if_else(
      major_group == "Not provided",
      NA_real_,
      value
    ),
    pc_valid = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3),
    pc_all = janitor::round_half_up(value/sum(value), 3)
  )

```

As at 31 March 2021, there were `r scales::comma(orientation_totals$value[orientation_totals$major_group == "LGBO"], 1)` who declared their sexual orientation as either lesbian, gay, bisexual or something else other than heterosexual/straight, this equates to `r scales::percent(orientation_totals$pc_valid[orientation_totals$major_group == "LGBO"], 0.1)` of those who declared their sexual orientation.[^1]

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="orientation-chart">}}
```{r orientation_chart}
orientation_plot_df <- orientation_grade %>%
  filter(responsibility_level == "All employees") %>%
  select(sexual_orientation, value) %>%
  mutate(
    fill_group = if_else(
      sexual_orientation == "Not declared" | sexual_orientation == "Not reported",
      "Not provided",
      "Provided"
    ),
    chart_label = paste0(sexual_orientation, " (", scales::comma(value, 1), ")"),
    chart_label = fct_reorder(chart_label, as.numeric(sexual_orientation))
  )

orientation_plot <- ggplot(orientation_plot_df, 
                           aes(x = fct_rev(chart_label), y = value, fill = fill_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Provided" = unname(govukhugo::govuk_colours["dark-blue"]),
      "Not provided" = unname(govukhugo::govuk_colours["mid-grey"])
    )) +
  theme_void() +
  coord_flip() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 12, hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  orientation_plot,
  width = 800,
  height = 150,
  alt_title = "Civil Service headcount by sexual orientation",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by sexual orientation",
    "After the bars for sexual orientations there is one bar to show those",
    "whose have chosen not to declare their sexual orientation and one bar to",
    "show those whose sexual orientation is not reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="orientation-table">}}
```{r orientation_table}

orientation_table <- orientation_grade %>%
  filter(responsibility_level == "All employees") %>%
  mutate(pc_orientation = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)) %>%
  select(sexual_orientation, value, pc_orientation) %>%
  arrange(sexual_orientation)

govukhugo::govuk_datatable(
  orientation_table,
  title = "Civil Service headcount by sexual orientation",
  col_names = c("Sexual orientation", "Headcount", "Percent"),
    search = FALSE
) %>%
  DT::formatRound(columns = 2, digits = 0) %>%
  DT::formatPercentage(columns = 3, digits = 1)

```
{{< /tab >}}
{{< /tabset >}}
:::
::::

The proportion of staff who declared their sexual orientation as either lesbian, gay, bisexual or something else other than heterosexual/straight is similar across responsibility levels.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide" ref="orientation-grade-chart">}}
```{r orientation_grade_chart}
orientation_grade_plot_df <- orientation_grade %>%
  filter(responsibility_level != "All employees" &
           responsibility_level != "Not reported" &
           major_group != "Not provided") %>%
  group_by(responsibility_level, major_group) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  mutate(pc_orientation = value/sum(value)) %>%
  ungroup() %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants"
    ),
    major_group = factor(major_group,
                         levels = c("Heterosexual", "LGBO"),
                         labels = c("Heterosexual/straight",
                                    "Lesbian, gay, bisexual or other")),
    pc_label = scales::percent(janitor::round_half_up(pc_orientation, 2), 1)
  )

orientation_grade_plot <- ggplot(
  orientation_grade_plot_df,
  aes(x = fct_rev(responsibility_level), y = pc_orientation, fill = fct_rev(major_group))
  ) +
  geom_col(position = "fill") +
  geom_text(aes(label = pc_label), 
            position = position_fill(vjust = 0.5),
            colour = govukhugo::govuk_colours["white"]) +
  scale_fill_manual(
    values = c(
      "Heterosexual/straight" = unname(govukhugo::govuk_colours["dark-blue"]),
      "Lesbian, gay, bisexual or other" =
        scales::gradient_n_pal(govukhugo::govuk_palette("dark-blue"))(c(0.5))
    ),
    name = "Sexual orientation"
  ) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(hjust = 1),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
    legend.position = "top",
    legend.direction = "horizontal"
  )

govukhugo::render_svg(
  orientation_grade_plot,
  width = 800,
  height = 200,
  alt_title = "Civil Service headcount by sexual orientation and responsibility level",
  alt_desc = paste(
    "A stacked bar chart showing Civil Service headcount by responsibility level",
    "and sexual orientation. Bars represent the proportion of civil servants in",
    "each responsibility level that have declared they are heterosexual/straight",
    "(the bar starting on the left) and those that have declared they are",
    "lesbian, gay, bisexual or another sexual orientation (the bar starting on",
    "the right). The bars are ordered by responsibility level with the most",
    "senior level at the top and administrative staff at the bottom. The chart",
    "does not include data for where either responsbility level or sexual",
    "orientation are not declared or reported."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide" ref="orientation-table-chart">}}
```{r orientation_grade_table}
orientation_grade_table <- orientation_grade %>%
  filter(responsibility_level != "All employees") %>%
  group_by(responsibility_level) %>%
  mutate(
    valid = if_else(major_group == "Not provided", NA_real_, value),
    pc_orientation = janitor::round_half_up(valid/sum(valid, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  mutate(
    responsibility_level = fct_relevel(
      responsibility_level, "Senior Civil Service Level", "Grades 6 and 7",
      "Senior and Higher Executive Officers", "Executive Officers",
      "Administrative Officers and Assistants", "Not reported"
    )
  ) %>%
  select(responsibility_level, sexual_orientation, value, pc_orientation) %>%
  arrange(responsibility_level, sexual_orientation)

govukhugo::govuk_datatable(
  orientation_grade_table,
  title = "Civil Service headcount by responsibility level and sexual orientation",
  col_names = c("Responsibility level", "Sexual orientation", "Headcount", "Percent"),
    page_length = 12,
    search = FALSE
) %>%
  DT::formatRound(columns = 3, digits = 0) %>%
  DT::formatPercentage(columns = 4, digits = 1)
```
{{< /tab >}}
{{< /tabset >}}
:::
::::

[^1]: The Civil Service Statistics record two types of missing value for sexual orientation. Where an individual has specifically chosen not to provide their sexual orientation, for example by selecting a "Prefer not to say" value, this is recorded as "Not declared". Where an individual has not provided any data on their sexual orientation, including not selecting a "Prefer not to say" value, this is recorded as "Not reported". There are `r scales::comma(orientation_totals$value[orientation_totals$major_group == "Not provided"], 1)` individuals whose sexual orientation is either not declared or not reported, which is `r scales::percent(orientation_totals$pc_all[orientation_totals$major_group == "Not provided"], 0.1)` of total headcount.
