---
title: "Civil servants by grade"
date: 2021-01-01
section: structure
weight: 100
summary: "Headline statistics of Civil Servants by grade"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

data_dir <- file.path(dirname(getwd()), "data")

acses2021 <- read_csv(file.path(data_dir, "acses2021_processed.csv"))

acses2021_grade_totals <- acses2021 %>%
  filter(source_table == "table_01" & 
           sex == "All employees" & 
           working_pattern == "All employees") %>%
  select(responsibility_level, value) %>%
  mutate(
    pc_all = value/max(value),
    valid = case_when(
      responsibility_level %in% c("Not reported", "All employees") ~ NA_real_,
      TRUE ~ value
    ),
    pc_valid = valid/sum(valid, na.rm = TRUE)
  )

acses2021_grade_numbers <- purrr::set_names(
  x  = scales::comma(acses2021_grade_totals$value),
  nm = acses2021_grade_totals$responsibility_level
)

acses2021_grade_pc_all <- purrr::set_names(
  x  = scales::percent(janitor::round_half_up(acses2021_grade_totals$pc_all, 3), 0.1),
  nm = acses2021_grade_totals$responsibility_level
)

acses2021_grade_pc_valid <- purrr::set_names(
  x  = scales::percent(janitor::round_half_up(acses2021_grade_totals$pc_valid, 3), 0.1),
  nm = acses2021_grade_totals$responsibility_level
)

acses2021_grade_org <- acses2021 %>%
  filter(source_table == "table_20") %>%
  select(organisation, responsibility_level, value, note) %>%
  mutate(
    valid = case_when(
      responsibility_level %in% c("Not reported", "All employees") ~ NA_real_,
      TRUE ~ value
    )
  )

acses2021_grade_org_max <- acses2021_grade_org %>%
  group_by(organisation) %>%
  mutate(
    pc_all = value/max(value, na.rm = TRUE),
    pc_valid = valid/sum(valid, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  filter(responsibility_level %in% c("Senior Civil Service Level", 
                                     "Administrative Officers and Assistants")) %>%
  left_join(
    acses2021_grade_org %>% 
      filter(responsibility_level == "All employees") %>%
      select(organisation, all_employees = value),
    by = "organisation"
  ) %>%
  anti_join(
    acses2021_grade_org %>% 
      filter(!is.na(note)) %>% 
      distinct(organisation),
    by = "organisation"
  ) %>%
  filter(value != 0 & organisation != "All employees" & all_employees >= 250) %>%
  group_by(responsibility_level) %>%
  filter(pc_valid == max(pc_valid) | valid == max(valid)) %>%
  mutate(
    marker = if_else(pc_valid == max(pc_valid), "percent", "number"),
    out_value = if_else(
      marker == "percent",
      scales::percent(janitor::round_half_up(pc_valid, 3), 0.1),
      scales::comma(valid)
    ),
    name = paste(responsibility_level, 
                   if_else(pc_valid == max(pc_valid), "percent", "number")),
  )

max_grade_values <- purrr::set_names(
  x  = acses2021_grade_org_max$out_value,
  nm = acses2021_grade_org_max$name
)

max_grade_orgs <- purrr::set_names(
  x  = acses2021_grade_org_max$organisation,
  nm = acses2021_grade_org_max$name
)

grade_org_out <- acses2021_grade_org %>%
  mutate(value = if_else(!is.na(note), NA_real_, value)) %>%
  select(organisation, responsibility_level, value) %>%
  pivot_wider(names_from = responsibility_level, values_from = value) %>%
  select(
    Organisation = organisation,
    all_of(c("Senior Civil Service Level", "Grades 6 and 7",
           "Senior and Higher Executive Officers", "Executive Officers", 
           "Administrative Officers and Assistants", "Not reported",
           "All employees"))
  ) %>%
  arrange(desc(`All employees`))

```

As at 31 March 2021 three in ten civil servants (`r acses2021_grade_pc_valid["Administrative Officers and Assistants"]`) were in an administrative grade[^1], compared to almost half of civil servants (47.1%) in 2011 being at an administrative grade. There are now `r acses2021_grade_numbers["Senior Civil Service Level"]` civil servants are at Senior Civil Service Level[^2], this equates to `r acses2021_grade_pc_valid["Senior Civil Service Level"]` of the Civil Service.

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide">}}
```{r grade_pyramid}

acses2021_grade_pyramid <- acses2021_grade_totals %>%
  drop_na(valid) %>%
  select(responsibility_level, value, pc_valid) %>%
  mutate(
    responsibility_level = factor(
      responsibility_level, 
      levels = c("Administrative Officers and Assistants",
                 "Executive Officers",
                 "Senior and Higher Executive Officers",
                 "Grades 6 and 7",
                 "Senior Civil Service Level"),
      ordered = TRUE),
    value_label = paste0(
      responsibility_level, " (", scales::comma(value), ", ", 
      scales::percent(janitor::round_half_up(pc_valid, 3), 0.1), ")"
    ),
    value_label = str_wrap(value_label, 30),
    bar_width = value/2
  ) %>%
  arrange(responsibility_level)

pyramid_plot <- ggplot(acses2021_grade_pyramid,
                       aes(x = responsibility_level, y = bar_width)) +
  geom_col(fill = govukhugo::govuk_colours["blue"]) +
  geom_col(aes(y = -bar_width), fill = govukhugo::govuk_colours["blue"]) +
  geom_text(aes(label = value_label), hjust = 0, nudge_y = 5000, size = 4) +
  scale_y_continuous(limits = c(-75000, 150000), expand = expansion()) +
  coord_flip() +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  pyramid_plot,
  width = 600,
  height = 250,
  alt_title = "Civil Service headcount by grade",
  alt_desc = paste(
    "A bar chart showing Civil Service headcount by grade. The bars are ordered",
    "by seniority level with the Senior Civil Service at the top and ",
    "administrative staff at the bottom so as to resemble a pyramid."
  ),
  caption = "alt_title"
)

```
{{< /tab >}}
{{< tab title="Table" heading="hide">}}
```{r grade_table}

acses2021_grade_pyramid %>%
  select(responsibility_level, value, pc_valid) %>%
  mutate(value = scales::comma(value),
         pc_valid = scales::percent(janitor::round_half_up(pc_valid, 3), 0.1)) %>%
  arrange(desc(responsibility_level)) %>%
  knitr::kable(
    format = "html", 
    align = "lrr",
    col.names = c("Grade/responsibility level", "Number", "Percent"),
    caption = "Civil Service headcount by grade"
  )
```
{{< /tab >}}
{{< /tabset >}}

:::
::::

## Grade by organisation

Organisations with large numbers of front-line staff delivering public services tend to have a higher proportion of junior grades. `r max_grade_orgs["Administrative Officers and Assistants percent"]` has the largest proportion of staff at an administrative grade (`r max_grade_values["Administrative Officers and Assistants percent"]`), while `r max_grade_orgs["Administrative Officers and Assistants number"]` has the largest number of staff at an administrative grade (`r max_grade_values["Administrative Officers and Assistants number"]`).

Organisations with more senior grades tend to have fewer employees delivering services directly to the public. `r max_grade_orgs["Senior Civil Service Level percent"]` has the largest proportion of staff at Senior Civil Service Level (`r max_grade_values["Senior Civil Service Level percent"]`), while `r max_grade_orgs["Senior Civil Service Level number"]` has the largest number of staff at Senior Civil Service Level (`r max_grade_values["Senior Civil Service Level number"]`)[^3].

::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
```{r org_dt, echo=FALSE}
govukhugo::govuk_datatable(
  data = grade_org_out,
  search = FALSE,
  options = list(
    columnDefs = list(
      list(defaultContent = "[c]", targets = 1:5)
    )
  )
) %>%
  DT::formatRound(columns = 2:8, digits = 0)
```
:::
::::



[^1]: Proportions and percentages are of those with a known grade. There were `r acses2021_grade_numbers["Not reported"]` staff whose grade was not reported, this is equivalent to `r acses2021_grade_pc_all["Not reported"]` of all employees.
[^2]: There are two measures for the most senior grades: the Senior Civil Service, and Senior Civil Service Level. See the [statistical notes](/about/#grading-systems-and-the-senior-civil-service) for more detail.
[^3]: This analysis excludes organisations with less than 250 employees overall.
