---
title: "Civil servants by organisation"
date: 2021-01-01
section: structure
weight: 100
summary: "Headline statistics of Civil Servants by organisation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

data_dir <- govukhugo::data_dir()

qpses_acses_2021 <- read_csv(
  file.path(data_dir, "qpses_acses_2021.csv")
)

group_lookup <- read_csv(
  file.path(data_dir, "dept_group_lookup.csv")
)

qpses_acses_totals_calc <- qpses_acses_2021 %>%
  filter(organisation == "All employees") %>%
  mutate(name = paste(data_source, value_type))

qpses_acses_totals <- purrr::set_names(
  x  = scales::comma(qpses_acses_totals_calc$value, 1),
  nm = qpses_acses_totals_calc$name
)

acses_hc <- qpses_acses_2021 %>%
  filter(value_type == "headcount" & 
           data_source == "ACSES" &
           organisation != "All employees") %>%
  left_join(
    group_lookup, by = "organisation"
  ) %>%
  add_count(dept_group, wt = value, name = "group_hc")

```

As at 31 March 2021, the Civil Service headcount[^1] stands at `r qpses_acses_totals["ACSES headcount"]`, an increase of 28,470[^2]. On a full-time equivalent (FTE) basis[^3] Civil Service employment stands at `r qpses_acses_totals["ACSES fte"]`.

The Civil Service Statistics reports on `r nrow(group_lookup)` organisations, which are further aggregated into `r length(unique(group_lookup$dept_group))` departmental groups[^4].


::::{.govuk-grid-row .data-card}
:::{.govuk-grid-column-full-width .data-display}
{{< tabset >}}
{{< tab title="Chart" heading="hide">}}
```{r group_chart}

group_plot_df <- acses_hc %>%
  distinct(dept_group, group_hc) %>%
  mutate(
    chart_group = case_when(
      group_hc < 1500 ~ "Other departments and agencies",
      grepl("Standards in Education", dept_group) ~ "Ofsted",
      TRUE ~ dept_group)
  ) %>%
  add_count(chart_group, wt = group_hc, name = "chart_group_hc") %>%
  add_count(chart_group, name = "chart_group_size") %>%
  mutate(
    chart_group_label = paste0(chart_group, 
                               " (", scales::comma(chart_group_hc), ")"),
    chart_group_label = fct_reorder(chart_group_label, group_hc, .fun = max)
  ) %>%
  distinct(chart_group_label, chart_group_hc)

group_plot <- ggplot(group_plot_df, aes(x = chart_group_label, y = chart_group_hc)) +
  geom_col(fill = govukhugo::govuk_colours["blue"]) +
  # geom_text(aes(label = chart_group_label),
  #           size = 3,
  #           colour = govukhugo::govuk_colours["govuk-secondary-text-colour"],
  #           hjust = 0, nudge_y = 1000) +
  scale_y_continuous(limits = c(0, 115000), expand = expansion()) +
  coord_flip() +
  theme_void() +
  theme(
    text = element_text(colour = govukhugo::govuk_colours["govuk-text-colour"]),
    axis.text.y = element_text(size = 10, hjust = 1, margin = margin(r = 3)),
    plot.background = element_rect(fill = "#f3f2f100", colour = NA),
  )

govukhugo::render_svg(
  group_plot,
  width = 600,
  height = 600,
  alt_title = "Civil Service headcount by departmental group",
  alt_desc = paste0(
    "A bar chart showing Civil Service headcount by departmental groups,",
    "arranged in descending order with the largest group (Work and Pensions)",
    "first. Departmental groups with less than 1,500 staff have been grouped",
    "together."
  ),
  caption = "alt_title"
)

```
{{< small >}}Departmental groups with a headcount of less than 1,500 staff have been grouped together.{{< /small >}}

{{< /tab >}}
{{< tab title="Departmental groups" heading="hide">}}
```{r group_table}

group_hcfte <- qpses_acses_2021 %>%
  filter(data_source == "ACSES" &
           organisation != "All employees") %>%
  left_join(
    group_lookup, by = "organisation"
  ) %>%
  pivot_wider(names_from = value_type, values_from = value) %>%
  group_by(dept_group) %>%
  summarise(across(c(headcount, fte), sum), .groups = "drop") %>%
  arrange(desc(headcount))

govukhugo::govuk_datatable(
    group_hcfte,
    title = "Civil Service employment by departmental group",
    col_names = c("Departmental group", "Headcount", "FTE"),
    page_length = 10,
    search = FALSE
  ) %>%
  DT::formatRound(columns = 2:3, digits = 0)

```
{{< /tab >}}
{{< tab title="Organisations" heading="hide">}}
```{r org_table}

org_hcfte <- qpses_acses_2021 %>%
  filter(data_source == "ACSES" &
           organisation != "All employees") %>%
  left_join(
    group_lookup, by = "organisation"
  ) %>%
  pivot_wider(names_from = value_type, values_from = value) %>%
  group_by(dept_group) %>%
  add_count(dept_group, wt = headcount, name = "group_hc") %>%
  mutate(dept_group = fct_reorder(dept_group, group_hc, .desc = TRUE)) %>%
  arrange(dept_group, desc(headcount)) %>%
  select(dept_group, organisation, headcount, fte)

govukhugo::govuk_datatable(
    org_hcfte,
    title = "Civil Service employment by organisation",
    col_names = c("Departmental group", "Organisation", "Headcount", "FTE"),
    page_length = 10,
    search = FALSE
  ) %>%
  DT::formatRound(columns = 3:4, digits = 0)

```
{{< /tab >}}
{{< /tabset >}}

:::
::::


[^1]: Add note about QPSES and ACSES.
[^2]: Figures and increase calculations here do not include the 19,310 Census 2021 temporary staff at UK Statistics Authority. See the [statistical notes](/about/#census-2021-temporary-staff) for more detail.
[^3]: Full-time equivalent (FTE) takes into account the hours worked by part-time employees. For example, an individual who works 4 days instead of 5 has an FTE value of 0.8. If an organisation had 200 staff (on a headcount basis) and 150 are employed full-time, 20 are employed for 4 days a week (0.8), 20 are employed for 3 days a week (0.6) and 10 are employed for 2 days a week (0.4) then the total FTE is 182, which would be reported as 180 (as figures are rounded to the nearest 10).
[^4]: Add note on organisational/departmental structure.
